---
title: "weeklyForecast"
output: 
  #rmarkdown::github_document
  html_document:
    toc: true
    number_sections: true
    keep_md: true
---



```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(tidyverse)
library(nlme)
library(RCurl)

rm(list = ls())
```


# Automate extracting 95% CI for prediction band, metrics, and plotting

Predict one day at a time, feedforward the prediction as lag case.rate. 

```{r forecastFunc}

forecastLME <- function(model, test) # provide model obj and test set
{
  for (i in 1:nrow(test))
  {
    
    pred <- predict(model, newdata = test[i,])
    test[i, 'pred'] <- pred
    
    # if second day or later, use pred case.rate as lag-1
    if ( i <= (nrow(test)-1) & as.Date(test[i+1,'date']) > cutOff+1)
    {
      test[i+1, 'casesTminus1.rate'] <- pred
    }
    
    # if third day or later, use pred case.rate as lag-2
    if (i <= (nrow(test)-2) & as.Date(test[i+2,'date']) > (cutOff + 2))
    {
      test[i+2, 'casesTminus2.rate'] <- pred
    }
  
  }
  
  return(test)
}

#forecast(model = case.rate.lme20201031
#         , test = test14
#         )

```


Extract 95% CI for prediction band. References from the web:

- https://stackoverflow.com/questions/14358811/extract-prediction-band-from-lme-fit and 
- http://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#predictions-andor-confidence-or-prediction-intervals-on-predictions

SE: This approach takes into account the uncertainty of the random effect in the prediction. 

SE2: This takes into account both random effect uncertainty and within-group error



``` {r resultsFunc}
#model <- case.rate.lme20201031

metricsFuncLME <- function(model)
{
  results7 <- forecastLME(model, test = test7)
  
  # create design mateix
  # [-2] drops response from formula
  Designmat <- model.matrix(formula(model)[-2], results7)
  
  # compute XVX′ to get the variance-covariance matrix of the predictions
  # extract the diagonal of this matrix to get variances of predictions
  predvar <- diag(Designmat %*% vcov(model) %*% t(Designmat)) 
  results7$SE <- sqrt(predvar) 
  results7$SE2 <- sqrt(predvar + model$sigma^2) # sigma is the estimated within-group error standard deviation
  
  
  results14 <- forecastLME(model, test = test14)
  
  Designmat <- model.matrix(formula(model)[-2], results14)
  predvar <- diag(Designmat %*% vcov(model) %*% t(Designmat)) 
  results14$SE <- sqrt(predvar) 
  results14$SE2 <- sqrt(predvar + model$sigma^2)


  # CALCULATE MAE, MAPE, CP
  perform7 <- results7 %>% 
  group_by(city) %>% 
  summarize(MAE = round(mean(abs(case.rate - pred)) , 2)
            , MAPE = round(mean(abs(case.rate - pred)/case.rate) , 2)
            , CP = round(sum(case.rate >= (pred - 1.96*SE2) & case.rate <= (pred + 1.96*SE2) ) / n() , 2)
            ) #%>%
  #as.data.frame() %>%
  #print()


  perform14 <- results14 %>% 
    group_by(city) %>% 
    summarize(MAE = round(mean(abs(case.rate - pred)) , 2)
              , MAPE = round(mean(abs(case.rate - pred)/case.rate) , 2)
              , CP = round(sum(case.rate >= (pred - 1.96*SE2) & case.rate <= (pred + 1.96*SE2) ) / n() , 2)
              ) #%>%
    #as.data.frame() %>%
    #print()
  
  metrics <- perform7 %>%
    left_join(perform14 
              , by = "city"
              , suffix = c("_7d", "_14d")
              ) #%>%
    #as.data.frame() %>%
    #print()
  
  return(list(performMetrics = metrics
              , results7 = results7
              , results14 = results14))
}




plotFuncLME <- function(model
                     , forecastLength # only 7 or 14 as numeric arg
                     )
{
  
  if(forecastLength == 7)
  { results = output$results7
    
    perform <- output$performMetrics %>% 
      select(city, contains("7d"))
  }
  
  if(forecastLength == 14)
  { results = output$results14
    
    perform <- output$performMetrics %>% 
      select(city, contains("14d")) 
  }
  
  perform <- setNames(perform, gsub("_.*", "", names(perform)))
  
  perform <- perform %>%
    mutate(MAPE = paste0("MAPE = ", MAPE)
                      , MAE = paste0("MAE = ", MAE)
                      , CP = paste0("CP = ", CP)
                      )
  
  
  Designmat <- model.matrix(formula(model)[-2], train)
  predvar <- diag(Designmat %*% vcov(model) %*% t(Designmat)) 
  
  plotDF <- train %>% 
    select(city, weekNum, case.rate) %>%
    mutate(city = as.character(city)
           , SE2 = sqrt(predvar + model$sigma^2)
           , upperCI.fitted = case.rate + 1.96*SE2
           , lowerCI.fitted = case.rate - 1.96*SE2
           , fitted = predict(model, data = train)
           ) %>%
    reshape::melt(id = c('city', 'weekNum')) %>%
    
    # bring into forecasted case rates
    rbind(results %>% 
            select(city, weekNum, case.rate, pred, SE2) %>% 
            #rename(case.rate = pred) %>% 
            mutate(upperCI.pred = pred + 1.96*SE2
                   , lowerCI.pred = pred - 1.96*SE2
                   ) %>%
            reshape2::melt(id = c('city', 'weekNum'))
          ) 
      
  
  p <- ggplot(plotDF 
              , aes(weekNum, value, group = variable, colour = variable)) +
    
    geom_line(data = filter(plotDF, variable == 'case.rate'), linetype = "solid", color = 'black', alpha = 0.6) + 
    geom_line(data = filter(plotDF, variable == 'fitted'), linetype = "solid", color = 'red') +
    geom_line(data = filter(plotDF, variable == 'pred'), linetype = "solid", color = 'blue') +
    
    # uncertainty band for fitted model
    geom_ribbon(data = filter(plotDF, variable == 'lowerCI.fitted')
      , aes(ymin = filter(plotDF, variable == 'lowerCI.fitted')$value
                    , ymax =  filter(plotDF, variable == 'upperCI.fitted')$value
                    )
                , alpha = 0.3
                , color = NA
                , fill = "red") +
    
    # prediction band for forecast
    geom_ribbon(data = filter(plotDF, variable == 'lowerCI.pred')
      , aes(ymin = filter(plotDF, variable == 'lowerCI.pred')$value
                    , ymax =  filter(plotDF, variable == 'upperCI.pred')$value
                    )
                , alpha = 0.3
                , color = NA
                , fill = "blue") +
    
    facet_wrap(. ~ city, ncol = 2) +
    ggtitle(paste0('Forecast ', forecastLength, ' Days')) +
    ylab('Case Rate per 100,000 People') +
    xlab('Days since March 16, 2020') +
    #ylim(-15,100) +
    
    # add performance metrics
    geom_label(data = perform 
               , aes(label = MAPE), 
              x = 0, y = max(abs(filter(plotDF, variable == 'case.rate')$value)), hjust="inward", vjust="inward",
              inherit.aes = FALSE) +
    geom_label(data = perform 
               , aes(label = MAE), 
              x = 0, y = max(abs(filter(plotDF, variable == 'case.rate')$value))*4/5, hjust="inward", vjust="inward",
              inherit.aes = FALSE) +
    geom_label(data = perform 
                , aes(label = CP), 
               x = 0, y = max(abs(filter(plotDF, variable == 'case.rate')$value))*2/5, hjust="inward", vjust="inward",
               inherit.aes = FALSE) +
     
    theme_classic() +
    theme(legend.position = 'none') 
    
  p
}

```


# Naive Multiple Regression

Automate some functions specific to LM object
```{r}
forecastLM <- function(model, test) # provide model obj and test set
{
  for (i in 1:nrow(test))
  {
    
    pred <- predict(model, newdata = test[i,], interval = "prediction", level = 0.95) %>% as.data.frame()
    test[i, 'pred'] <- pred$fit
    test[i, 'lowerCI'] <- pred$lwr
    test[i, 'upperCI'] <- pred$upr
    
    
    # if second day or later, use pred case.rate as lag-1
    if ( i <= (nrow(test)-1) & as.Date(test[i+1,'date']) > cutOff+1)
    {
      test[i+1, 'casesTminus1.rate'] <- pred$fit
    }
    
    # if third day or later, use pred case.rate as lag-2
    if (i <= (nrow(test)-2) & as.Date(test[i+2,'date']) > (cutOff + 2))
    {
      test[i+2, 'casesTminus2.rate'] <- pred$fit
    }
  
  }
  
  return(test)
}

# 
# 
# temp <- forecastLM(model = case.rate.lm
#         , test = test14
#         )


metricsFuncLM <- function(model)
{
  results7 <- forecastLM(model, test = test7)
  
  results14 <- forecastLM(model, test = test14)
  
  metrics <- results7 %>% 
    group_by(city) %>% 
    summarize(MAE = round(mean(abs(case.rate - pred)) , 2)
              , MAPE = round(mean(abs(case.rate - pred)/case.rate) , 2)
              , CP = round(sum(case.rate >= lowerCI & case.rate <= upperCI ) / n() , 2)
              ) %>%
    left_join( results14 %>% 
                 group_by(city) %>% 
                 summarize(MAE = round(mean(abs(case.rate - pred)) , 2)
                           , MAPE = round(mean(abs(case.rate - pred)/case.rate) , 2)
                           , CP = round(sum(case.rate >= lowerCI & case.rate <= upperCI ) / n() , 2)
                           )
               , by = "city"
               , suffix = c("_7d", "_14d")
               ) 
  
  return(list(performMetrics = metrics
              , results7 = results7
              , results14 = results14))
}




plotFuncLM <- function(model, forecastLength)
{
  
  
  if(forecastLength == 7)
  { results = output$results7
    
    metrics <- output$performMetrics %>% 
      select(city, contains("7d"))
  }
  
  if(forecastLength == 14)
  { results = output$results14
    
    metrics <- output$performMetrics %>% 
      select(city, contains("14d")) 
  }
  
  metrics <- setNames(metrics, gsub("_.*", "", names(metrics)))

  pred <- predict(model, newdata = train, interval = "prediction", level = 0.95)

  plotDF <- train %>% 
    select(city, day031620, case.rate) %>%
    mutate(city = as.character(city)) %>%
    cbind(pred) %>%
    rename(fitted = fit
           , lowerCI.fitted = lwr
           , upperCI.fitted = upr
           ) %>%
    reshape::melt(id = c('city', 'day031620')) %>%
    
    # bring into forecasted case rates
    rbind(results %>% 
            select(city, day031620, case.rate, pred, upperCI, lowerCI) %>% 
            rename(upperCI.pred = upperCI
                   , lowerCI.pred = lowerCI
                   ) %>%
            reshape2::melt(id = c('city', 'day031620'))
          ) 
      
  
  p <- ggplot(plotDF 
              , aes(day031620, value, group = variable, colour = variable)) +
    
    geom_line(data = filter(plotDF, variable == 'case.rate'), linetype = "solid", color = 'black', alpha = 0.6) + 
    geom_line(data = filter(plotDF, variable == 'fitted'), linetype = "solid", color = 'red') +
    geom_line(data = filter(plotDF, variable == 'pred'), linetype = "solid", color = 'blue') +
    
    # uncertainty band for fitted model
    geom_ribbon(data = filter(plotDF, variable == 'lowerCI.fitted')
      , aes(ymin = filter(plotDF, variable == 'lowerCI.fitted')$value
                    , ymax =  filter(plotDF, variable == 'upperCI.fitted')$value
                    )
                , alpha = 0.3, color = NA
                , fill = "red") +
    
    # prediction band for forecast
    geom_ribbon(data = filter(plotDF, variable == 'lowerCI.pred')
      , aes(ymin = filter(plotDF, variable == 'lowerCI.pred')$value
                    , ymax =  filter(plotDF, variable == 'upperCI.pred')$value
                    )
                , alpha = 0.3, color = NA
                , fill = "blue") +
    
    facet_wrap(. ~ city, ncol = 2) +
    ggtitle('Forecast 7 Days') +
    ylab('Case Rate per 100,000 People') +
    xlab('Days since March 16, 2020') +
    #ylim(-15,100) +
    
    # add performance metrics
    geom_label(data = metrics %>% mutate(MAPE = paste0("MAPE = ", MAPE))
               , aes(label = MAPE), 
              x = 0, y = max(abs(filter(plotDF, variable == 'case.rate')$value)), hjust="inward", vjust="inward",
              inherit.aes = FALSE) +
    geom_label(data = metrics %>% mutate(MAE = paste0("MAE = ", MAE))
               , aes(label = MAE),
              x = 0, y = max(abs(filter(plotDF, variable == 'case.rate')$value))*4/5, hjust="inward", vjust="inward",
              inherit.aes = FALSE) +
    geom_label(data = metrics %>% mutate(CP = paste0("CP = ", CP))
                , aes(label = CP),
               x = 0, y = max(abs(filter(plotDF, variable == 'case.rate')$value))*3/5, hjust="inward", vjust="inward",
               inherit.aes = FALSE) +

    theme_classic() +
    theme(legend.position = 'none') 
  
  
  p
}

```


# Mar-Nov
``` {r}

weeklyCases <- read.csv('https://raw.githubusercontent.com/kentranz/socialMobilityCOVID/master/data/weeklyCases.csv')

cutOff <- as.Date("2020-11-29")

train <- weeklyCases %>% 
  filter(weekNum > 3 & as.Date(date) < cutOff)


test7 <- weeklyCases %>% 
   filter(weekNum > 3 & as.Date(date) >= cutOff & weekNum < 41)

  
test14 <- weeklyCases %>% 
  filter(weekNum > 3 & as.Date(date) >= cutOff) %>%
  
  # retained lag case rates for first week only, zero out the rest
  mutate(
    casesTminus2.rate = case_when(weekNum == 41 ~ 0, TRUE ~ casesTminus2.rate)
  )


nrow(train)
nrow(test7)
nrow(test14)
```



## LME
``` {r}

lme <- 
  lme(fixed = case.rate ~ casesTminus1.rate + casesTminus2.rate,
      random = ~ casesTminus1.rate + casesTminus2.rate - 1 | city,
      data = train
      )

summary(lme)

```


``` {r fig.width=10, fig.height=30}

output <- metricsFuncLME(model = lme)

allMetrics <- output$performMetrics %>%
      rename_at(vars(-city), function(x) {paste0(x, "_lme")})


```

## LME weighted

```{r }
lmeWt <- 
  lme(fixed = case.rate ~ casesTminus1.rate + casesTminus2.rate
      , random = ~ casesTminus1.rate + casesTminus2.rate - 1 | city
      , data = train
      , weights = varPower(form = ~ fitted(.))
      , control = lmeControl(maxIter = 1e4) 
      )

summary(lmeWt)

output <- metricsFuncLME(model = lmeWt)

allMetrics <- allMetrics %>% left_join(
    output$performMetrics %>%
      rename_at(vars(-city), function(x) {paste0(x, "_lmeWt")})
    , by = "city"
    )

```



## Multiple Regression
```{r}
lm <- 
  lm( case.rate ~ casesTminus1.rate + casesTminus2.rate
       , data = train
       )

summary(lm)
```
```{r fig.width=10, fig.height=30}

output <- metricsFuncLM(model = lm)

allMetrics <- allMetrics %>% left_join(
    output$performMetrics %>%
      rename_at(vars(-city), function(x) {paste0(x, "_lm")})
    , by = "city"
    )
```

## LME + Testing Near Me

```{r }
lmeTest <- 
  lme(fixed = case.rate ~ casesTminus1.rate + casesTminus2.rate + testingNearMe
      , random = ~ casesTminus1.rate + casesTminus2.rate - 1 | city
      , data = train
      )

summary(lmeTest)

output <- metricsFuncLME(model = lmeTest)

allMetrics <- allMetrics %>% left_join(
    output$performMetrics %>%
      rename_at(vars(-city), function(x) {paste0(x, "_lmeTest")})
    , by = "city"
    )

```

## Multiple Regression + Testing Near Me
```{r}
lmTest <- 
  lm( case.rate ~ casesTminus1.rate + casesTminus2.rate + testingNearMe
       , data = train
       )

summary(lmTest)
```
```{r fig.width=10, fig.height=30}

output <- metricsFuncLM(model = lmTest)

allMetrics <- allMetrics %>% left_join(
    output$performMetrics %>%
      rename_at(vars(-city), function(x) {paste0(x, "_lmTest")})
    , by = "city"
    )
```


## Log-transformed

```{r}

train <- train %>% 
  mutate(case.rate = log(case.rate)
         , casesTminus1.rate = case_when(casesTminus1.rate != 0 ~ log(casesTminus1.rate)
                                         , TRUE ~ 0.000000001)
         , casesTminus2.rate = case_when(casesTminus2.rate != 0 ~ log(casesTminus2.rate)
                                         , TRUE ~ 0.000000001)
         )

test7 <- test7  %>% 
  mutate(case.rate = log(case.rate)
         , casesTminus1.rate = case_when(casesTminus1.rate != 0 ~ log(casesTminus1.rate)
                                         , TRUE ~ 0.000000001)
         , casesTminus2.rate = case_when(casesTminus2.rate != 0 ~ log(casesTminus2.rate)
                                         , TRUE ~ 0.000000001)
         )
         

test14 <- test14  %>% 
  mutate(case.rate = log(case.rate)
         , casesTminus1.rate = case_when(casesTminus1.rate != 0 ~ log(casesTminus1.rate)
                                         , TRUE ~ 0.000000001)
         , casesTminus2.rate = case_when(casesTminus2.rate != 0 ~ log(casesTminus2.rate)
                                         , TRUE ~ 0.000000001)
         )
```
## LME-Log
NOT CONVERGING - TBD

<!-- ```{r } -->
<!-- lmeLog <- -->
<!--   lme(fixed = case.rate ~ casesTminus1.rate + casesTminus2.rate -->
<!--       , random = ~ casesTminus1.rate + casesTminus2.rate - 1 | city -->
<!--       , data = train -->
<!--       , control = lmeControl(maxIter = 1e8) -->
<!--       ) -->

<!-- summary(lmeLog) -->

<!-- output <- metricsFuncLME(model = lmeLog) -->

<!-- allMetrics <- allMetrics %>% left_join( -->
<!--     output$performMetrics %>% -->
<!--       rename_at(vars(-city), function(x) {paste0(x, "_lmeLog")}) -->
<!--     , by = "city" -->
<!--     ) -->

<!-- ``` -->


## LME-Log weighted residual
```{r}
lmeLogWt <- 
  lme(fixed = case.rate ~ casesTminus1.rate + casesTminus2.rate
      , random = ~ casesTminus1.rate + casesTminus2.rate - 1 | city
      , data = train
      , weights = varPower(form = ~ fitted(.))
      , control = lmeControl(maxIter = 1e4) 
      )

summary(lmeLogWt)
```
``` {r fig.width=10, fig.height=30}

output <- metricsFuncLME(model = lmeLogWt)

allMetrics <- allMetrics %>% left_join(
    output$performMetrics %>%
      rename_at(vars(-city), function(x) {paste0(x, "_lmeLogWt")})
    , by = "city"
    )
```


## Multiple Regression
```{r}
lmLog <- 
  lm( case.rate ~ casesTminus1.rate + casesTminus2.rate
       , data = train
       )

summary(lmLog)
```
```{r fig.width=10, fig.height=30}

output <- metricsFuncLM(model = lmLog)

allMetrics <- allMetrics %>% left_join(
    output$performMetrics %>%
      rename_at(vars(-city), function(x) {paste0(x, "_lmLog")})
    , by = "city"
    )
```

## LME-Log + Testing Near Me

```{r }
lmeLogTest <- 
  lme(fixed = case.rate ~ casesTminus1.rate + casesTminus2.rate + testingNearMe
      , random = ~ casesTminus1.rate + casesTminus2.rate - 1 | city
      , data = train
      )

summary(lmeLogTest)

output <- metricsFuncLME(model = lmeLogTest)

allMetrics <- allMetrics %>% left_join(
    output$performMetrics %>%
      rename_at(vars(-city), function(x) {paste0(x, "_lmeLogTest")})
    , by = "city"
    )

```



## LME-Log weighted residual + Testing Near Me
```{r}
lmeLogWtTest <- 
  lme(fixed = case.rate ~ casesTminus1.rate + casesTminus2.rate + testingNearMe
      , random = ~ casesTminus1.rate + casesTminus2.rate - 1 | city
      , data = train
      , weights = varPower(form = ~ fitted(.))
      , control = lmeControl(maxIter = 1e4) 
      )

summary(lmeLogWtTest)
```
``` {r fig.width=10, fig.height=30}

output <- metricsFuncLME(model = lmeLogWtTest)

allMetrics <- allMetrics %>% left_join(
    output$performMetrics %>%
      rename_at(vars(-city), function(x) {paste0(x, "_lmeLogWtTest")})
    , by = "city"
    )
```

## Multiple Regression
```{r}
lmLogTest <- 
  lm( case.rate ~ casesTminus1.rate + casesTminus2.rate + testingNearMe
       , data = train
       )

summary(lmLogTest)
```
```{r fig.width=10, fig.height=30}

output <- metricsFuncLM(model = lmLogTest)

allMetrics <- allMetrics %>% left_join(
    output$performMetrics %>%
      rename_at(vars(-city), function(x) {paste0(x, "_lmLogTest")})
    , by = "city"
    )
```

## Metrics for Mar-Nov
``` {r}

allMetrics %>% select(city, starts_with("MAPE_7d")) %>%
  arrange(as.character(city)) %>%
  as.data.frame() %>%
  format(scientific = F) %>%
  print()

```

``` {r}

allMetrics %>% select(city, starts_with("MAE_7d")) %>%
  arrange(as.character(city)) %>% 
  as.data.frame() %>%
  print()

```

``` {r}

allMetrics %>% select(city, starts_with("CP_7d")) %>%
  arrange(as.character(city)) %>%
  as.data.frame() %>%
  print()

```



``` {r}

allMetrics %>% select(city, starts_with("MAPE_14d")) %>%
  arrange(as.character(city)) %>%
  as.data.frame() %>%
  format(scientific = F) %>%
  print()

```

``` {r}

allMetrics %>% select(city, starts_with("MAE_14d")) %>%
  arrange(as.character(city)) %>% 
  as.data.frame() %>%
  print()

```

``` {r}

allMetrics %>% select(city, starts_with("CP_14d")) %>%
  arrange(as.character(city)) %>%
  as.data.frame() %>%
  print()

```








# Mar-Aug
``` {r trainTest}

cutOff <- as.Date("2020-08-31")

train <- weeklyCases %>% 
  filter(weekNum > 3 & as.Date(date) < cutOff)


test7 <- weeklyCases %>% 
   filter(weekNum > 3 & as.Date(date) >= cutOff & weekNum <= 29)

  
test14 <- weeklyCases %>% 
  filter(weekNum > 3 & as.Date(date) >= cutOff) %>%
  
  # retained lag case rates for first week only, zero out the rest
  mutate(
    casesTminus2.rate = case_when(weekNum == 29 ~ 0, TRUE ~ casesTminus2.rate)
  )


nrow(train)
nrow(test7)
nrow(test14)
```



## LME
``` {r}

lme <- 
  lme(fixed = case.rate ~ casesTminus1.rate + casesTminus2.rate,
      random = ~ casesTminus1.rate + casesTminus2.rate - 1 | city,
      data = train
      )

summary(lme)

```


``` {r fig.width=10, fig.height=30}

output <- metricsFuncLME(model = lme)

allMetrics <- output$performMetrics %>%
      rename_at(vars(-city), function(x) {paste0(x, "_lme")})


```

## LME weighted

```{r }
lmeWt <- 
  lme(fixed = case.rate ~ casesTminus1.rate + casesTminus2.rate
      , random = ~ casesTminus1.rate + casesTminus2.rate - 1 | city
      , data = train
      , weights = varPower(form = ~ fitted(.))
      , control = lmeControl(maxIter = 1e4) 
      )

summary(lmeWt)

output <- metricsFuncLME(model = lmeWt)

allMetrics <- allMetrics %>% left_join(
    output$performMetrics %>%
      rename_at(vars(-city), function(x) {paste0(x, "_lmeWt")})
    , by = "city"
    )

```



## Multiple Regression
```{r}
lm <- 
  lm( case.rate ~ casesTminus1.rate + casesTminus2.rate
       , data = train
       )

summary(lm)
```
```{r fig.width=10, fig.height=30}

output <- metricsFuncLM(model = lm)

allMetrics <- allMetrics %>% left_join(
    output$performMetrics %>%
      rename_at(vars(-city), function(x) {paste0(x, "_lm")})
    , by = "city"
    )
```

## LME + Testing Near Me

```{r }
lmeTest <- 
  lme(fixed = case.rate ~ casesTminus1.rate + casesTminus2.rate + testingNearMe
      , random = ~ casesTminus1.rate + casesTminus2.rate - 1 | city
      , data = train
      )

summary(lmeTest)

output <- metricsFuncLME(model = lmeTest)

allMetrics <- allMetrics %>% left_join(
    output$performMetrics %>%
      rename_at(vars(-city), function(x) {paste0(x, "_lmeTest")})
    , by = "city"
    )

```

## Multiple Regression + Testing Near Me
```{r}
lmTest <- 
  lm( case.rate ~ casesTminus1.rate + casesTminus2.rate + testingNearMe
       , data = train
       )

summary(lmTest)
```
```{r fig.width=10, fig.height=30}

output <- metricsFuncLM(model = lmTest)

allMetrics <- allMetrics %>% left_join(
    output$performMetrics %>%
      rename_at(vars(-city), function(x) {paste0(x, "_lmTest")})
    , by = "city"
    )
```


## Log-transformed

```{r}

train <- train %>% 
  mutate(case.rate = log(case.rate)
         , casesTminus1.rate = case_when(casesTminus1.rate != 0 ~ log(casesTminus1.rate)
                                         , TRUE ~ 0.000000001)
         , casesTminus2.rate = case_when(casesTminus2.rate != 0 ~ log(casesTminus2.rate)
                                         , TRUE ~ 0.000000001)
         )

test7 <- test7  %>% 
  mutate(case.rate = log(case.rate)
         , casesTminus1.rate = case_when(casesTminus1.rate != 0 ~ log(casesTminus1.rate)
                                         , TRUE ~ 0.000000001)
         , casesTminus2.rate = case_when(casesTminus2.rate != 0 ~ log(casesTminus2.rate)
                                         , TRUE ~ 0.000000001)
         )
         

test14 <- test14  %>% 
  mutate(case.rate = log(case.rate)
         , casesTminus1.rate = case_when(casesTminus1.rate != 0 ~ log(casesTminus1.rate)
                                         , TRUE ~ 0.000000001)
         , casesTminus2.rate = case_when(casesTminus2.rate != 0 ~ log(casesTminus2.rate)
                                         , TRUE ~ 0.000000001)
         )
```
## LME-Log
NOT CONVERGING - TBD

<!-- ```{r } -->
<!-- lmeLog <- -->
<!--   lme(fixed = case.rate ~ casesTminus1.rate + casesTminus2.rate -->
<!--       , random = ~ casesTminus1.rate + casesTminus2.rate - 1 | city -->
<!--       , data = train -->
<!--       , control = lmeControl(maxIter = 1e8) -->
<!--       ) -->

<!-- summary(lmeLog) -->

<!-- output <- metricsFuncLME(model = lmeLog) -->

<!-- allMetrics <- allMetrics %>% left_join( -->
<!--     output$performMetrics %>% -->
<!--       rename_at(vars(-city), function(x) {paste0(x, "_lmeLog")}) -->
<!--     , by = "city" -->
<!--     ) -->

<!-- ``` -->


## LME-Log weighted residual
```{r}
lmeLogWt <- 
  lme(fixed = case.rate ~ casesTminus1.rate + casesTminus2.rate
      , random = ~ casesTminus1.rate + casesTminus2.rate - 1 | city
      , data = train
      , weights = varPower(form = ~ fitted(.))
      , control = lmeControl(maxIter = 1e4) 
      )

summary(lmeLogWt)
```
``` {r fig.width=10, fig.height=30}

output <- metricsFuncLME(model = lmeLogWt)

allMetrics <- allMetrics %>% left_join(
    output$performMetrics %>%
      rename_at(vars(-city), function(x) {paste0(x, "_lmeLogWt")})
    , by = "city"
    )
```


## Multiple Regression
```{r}
lmLog <- 
  lm( case.rate ~ casesTminus1.rate + casesTminus2.rate
       , data = train
       )

summary(lmLog)
```
```{r fig.width=10, fig.height=30}

output <- metricsFuncLM(model = lmLog)

allMetrics <- allMetrics %>% left_join(
    output$performMetrics %>%
      rename_at(vars(-city), function(x) {paste0(x, "_lmLog")})
    , by = "city"
    )
```

## LME-Log + Testing Near Me
NOT CONVERGING

<!-- ```{r } -->
<!-- lmeLogTest <-  -->
<!--   lme(fixed = case.rate ~ casesTminus1.rate + casesTminus2.rate + testingNearMe -->
<!--       , random = ~ casesTminus1.rate + casesTminus2.rate - 1 | city -->
<!--       , data = train -->
<!--       ) -->

<!-- summary(lmeLogTest) -->

<!-- output <- metricsFuncLME(model = lmeLogTest) -->

<!-- allMetrics <- allMetrics %>% left_join( -->
<!--     output$performMetrics %>% -->
<!--       rename_at(vars(-city), function(x) {paste0(x, "_lmeLogTest")}) -->
<!--     , by = "city" -->
<!--     ) -->

<!-- ``` -->



## LME-Log weighted residual + Testing Near Me
```{r}
lmeLogWtTest <- 
  lme(fixed = case.rate ~ casesTminus1.rate + casesTminus2.rate + testingNearMe
      , random = ~ casesTminus1.rate + casesTminus2.rate - 1 | city
      , data = train
      , weights = varPower(form = ~ fitted(.))
      , control = lmeControl(maxIter = 1e4) 
      )

summary(lmeLogWtTest)
```
``` {r fig.width=10, fig.height=30}

output <- metricsFuncLME(model = lmeLogWtTest)

allMetrics <- allMetrics %>% left_join(
    output$performMetrics %>%
      rename_at(vars(-city), function(x) {paste0(x, "_lmeLogWtTest")})
    , by = "city"
    )
```

## Multiple Regression
```{r}
lmLogTest <- 
  lm( case.rate ~ casesTminus1.rate + casesTminus2.rate + testingNearMe
       , data = train
       )

summary(lmLogTest)
```
```{r fig.width=10, fig.height=30}

output <- metricsFuncLM(model = lmLogTest)

allMetrics <- allMetrics %>% left_join(
    output$performMetrics %>%
      rename_at(vars(-city), function(x) {paste0(x, "_lmLogTest")})
    , by = "city"
    )
```

## Metrics for Mar-Aug
``` {r}

allMetrics %>% select(city, starts_with("MAPE_7d")) %>%
  arrange(as.character(city)) %>%
  as.data.frame() %>%
  format(scientific = F) %>%
  print()

```

``` {r}

allMetrics %>% select(city, starts_with("MAE_7d")) %>%
  arrange(as.character(city)) %>% 
  as.data.frame() %>%
  print()

```

``` {r}

allMetrics %>% select(city, starts_with("CP_7d")) %>%
  arrange(as.character(city)) %>%
  as.data.frame() %>%
  print()

```



``` {r}

allMetrics %>% select(city, starts_with("MAPE_14d")) %>%
  arrange(as.character(city)) %>%
  as.data.frame() %>%
  format(scientific = F) %>%
  print()

```

``` {r}

allMetrics %>% select(city, starts_with("MAE_14d")) %>%
  arrange(as.character(city)) %>% 
  as.data.frame() %>%
  print()

```

``` {r}

allMetrics %>% select(city, starts_with("CP_14d")) %>%
  arrange(as.character(city)) %>%
  as.data.frame() %>%
  print()

```
